---
title: "HW2"
author: "Xuesen Zhao"
date: "2022-10-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
library(tidyverse)
library(readxl)
```


```{r}
Transit = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>% 
  select(line:entry,exit_only, vending,  
    ada) %>%
  mutate(entry = ifelse(entry=="YES",TRUE,FALSE)) %>%
  mutate(vending = ifelse(vending=="YES",TRUE,FALSE)) 
```

The original dataset contains information about the entrance and exit for each listed subway station in NYC and it had 1868 observations and 32 rows. The column names were formatted and coerced into lower case, and 20 variables were selected from the original dataset, including line, station name, station latitude and longitude, whether a route (1-11) stops at that station, entrance type, entry or exit, vending, and ADA compliance. The resulting dataset now contains 1868 rows and 20 columns. 

Yet, the dataset is not completely clean. There are many rows that contains missing values, route number should be a separate variable. 

```{r}
# distinct station
distinct(Transit, station_name,line,.keep_all = TRUE) 
# ADA compliant
Transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
# Entry = TRUE, AND vending = FALSE
Transit %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

There are 465 stations that have a distinct combination of station name and line. 84 Stations are ada compliant. There is no station that has entrance/exits without vending but allow entrance. 

```{r}
Transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

Transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 60 distinct stations that serve the A train, and 17 of them are ADA compliant. 

## Question 2

```{r}
# The row with pic was excluded
Mr_wheel = read_excel("./data/Trash_Wheel.xlsx",sheet = 1, range = "A2:N549") %>%
  janitor::clean_names() %>% 
  mutate(Type="Mr",sports_balls=as.numeric(round(sports_balls,0)))

# The row with pic was excluded, 
Professor_wheel = read_excel("./data/Trash_Wheel.xlsx",sheet = 2, range = "A2:M96") %>%
  janitor::clean_names()%>%
  mutate(Type="Professor")

```

The first row in the original excel sheet was excluded because it contained only a picture. The resulting dataframe Mr_wheel, was generated based on the data from sheet 1, has `r nrow(Mr_wheel)` rows and `r ncol(Mr_wheel)` columns. It included variables such as the amount of plastic bottles, polystyrene, sports balls, grocery bags and other different types of trash removed by the wheel. For instance, the mean value of sports balls collected among these dumpster sites was `r mean(Mr_wheel$sports_ball)`. The Professor_wheel dataframe was generated based on the second sheet of the original excel file, and the first row which contained a picture was also skipped. The resulting dataframe contains `r nrow(Professor_wheel)` rows and`r ncol(Professor_wheel)` columns, including all variables listed in Mr_wheel except the number of sports balls collected. A type variable was added to each of the dataframe to indicate the type of trash wheel. 

```{r}
# The larger df Trash_wheel(x) is left_joined to the Professor trash wheel (y)
combined_wheel = left_join(
  Mr_wheel, Professor_wheel, by="dumpster"
)
combined_wheel %>%
  filter(Type.y=="Professor") %>%
  pull(weight_tons.y)%>%
  sum

combined_wheel %>%
  filter(Type.x=="Mr",year.x==2020)%>%
  pull(sports_balls)%>%
  sum
```

The Mr_wheel and Professor_wheel data sets were combined, based on the dumpster variable, to generate single dataframe that has `r nrow(combined_wheel)` rows and `r ncol(combined_wheel)` columns. The total weight of trash collected by Professor Trash Wheel was 190.12 tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856. 

## Question 3

```{r}
pol_month = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year","month","day"),sep = "-") %>%
  filter(prez_gop!=2)%>%
  mutate(
    year = as.numeric(year),
    month=month.abb[as.numeric(month)],
    president = ifelse(prez_gop==1,"gop","dem"))%>%
  select(-day,-prez_gop,-prez_dem)
  
```


```{r}
snp = read_csv("./data/snp.csv")%>%
  janitor::clean_names() %>%
  separate(date, into = c("month","day","year"),sep = "/") %>% 
  mutate(
    year = as.numeric(year), 
    year=ifelse(year<=20,2000+year,1900+year),
    month=month.abb[as.numeric(month)],
        ) %>%
  select(year,month,close,-day)
```

```{r}
unemploy = read_csv("./data/unemployment.csv") %>%
  janitor::clean_names()%>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment percentage"
  ) %>% 
  mutate(month=str_to_title(month))
```

The pol_month dataset contains information on the number of national politicians who are democratic or republican from 1947 to 2015. The cleaned dataset has `r nrow(pol_month)` rows and `r ncol(pol_month)` observations, including the variables year, month, and president (indicator of if the president is a republican or democratic), and the number of republican or democratic governors (gov), senators (sen), and representative (rep) on the same associated date. It does not make sense for a president to be both a republican and a democrat, so the rows with a value of 2 for prez_gop were omitted. 


A month and year variable with the same format were also created in the other two data frames as key variable for merging them together. The snp dataset contains `r nrow(snp)` rows and `r ncol(snp)` variables, including year (with an range of `r range(snp$year)`), month and the closing values of the S&P stock index on the associated date. The unemploy date contains `r nrow(unemploy)` rows and `r ncol(unemploy)` columns. It includes the unemployment percentage of all months with a year range of `r range(unemploy$year)`. 

```{r}
merge1 = left_join(pol_month,snp,by=c("year","month"))
merged_final = left_join(merge1, unemploy, by=c("year","month"))
```

The final merged dataframe of all three contains `r nrow(merged_final)` rows and `r ncol(merged_final)` columns, including all the variables from the three dataframe. The range of year for the merged data set is `r range(merged_final$year)`. 